Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-05-24 paper-amia-2018/manuscript.tex
update manuscript

\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns. When we predict the ``new segment'' point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior. Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:

@@

As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection. KNN also shows superior performance in all performance metrics for ``same segment'' class and weighted average over ``new segment'' and ``same segment'' classes. NB demonstrates the lowest performance among all models in terms of all performance metrics. On the other hand, SVM has the highest recall of 0.673 when only lexical features are used to identify ``new segment''. Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected since 96.71\% instances belong to ``same segment'' categories. For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in same segment identification compared to new segment detection. When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for new segment identification, 0.1\% higher precision than LSTM for same segment identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification. \\

@@

Similar to baseline performance, classification performance is remarkably higher for same segment class compared to new segment class (Table~\ref{tab:result_not_boundary}). The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class. We observed that precision increases by 0.4\% and 0.2\%, 0.6\% and 0.2\% and 0.6\% and 0.2\% for SVM and KNN; recall increases by 0\% and -0.1\%, 0.2\% and -0.1\% and 0.1\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.2\% and 0.1\%, 0.4\% and 0.1\% and 0.3\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively. Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying same segment, respectively. \\

>>>>

\textbf{Recurrent Neural Networks}: RNN is a neural network architecture designed to capture sequential patterns. When we predict the new segment point using RNNs, particular combinations of words in a sequence will indicate the change of MI behavior. Long Short Term Memory (LSTM) networks\cite{hochreiter1997long} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. GRU\cite{chung2014empirical} is a variant of LSTM mathematically represented by the following formula:

@@

As follows from Table~\ref{tab:result_base}, KNN outperforms all other models in terms of precision and F1-measure achieving 0.808 precision with 0.728 F1-measure for new segment detection. KNN also shows superior performance in all performance metrics for ``same segment'' class and weighted average over ``new segment'' and ``same segment'' classes. NB demonstrates the lowest performance among all models in terms of all performance metrics. On the other hand, SVM has the highest recall of 0.673 when only lexical features are used to identify ``new segment''. Results indicate that performance of classifiers is remarkably higher for ``same segment'' class compared to ``new segment'' class, which is expected since 96.71\% instances belong to ``same segment'' categories. For example, KNN achieves 22.40\%, 50.07\% and 36.26\% higher precision, recall and F1-measure, respectively, in ``same segment'' identification compared to ``new segment'' detection. When lexical features are used in combination with punctuation mark, GRU demonstrates the best performance in all cases showing 10.68\% higher recall and 3.78\%  higher F1-measure than LSTM for ``new segment'' identification, 0.1\% higher precision than LSTM for ``same segment'' identification and 0.2\% higher recall and 0.2\%  higher F1-measure than LSTM for overall classification. \\

@@

Similar to baseline performance, classification performance is remarkably higher for ``same segment'' class compared to ``new segment'' class (Table~\ref{tab:result_not_boundary}). The impact of punctuation and topic model-based features is also consistent with ``new segment'' classification while improvements are not as significant as ``new segment'' class. We observed that precision increases by 0.4\% and 0.2\%, 0.6\% and 0.2\% and 0.6\% and 0.2\% for SVM and KNN; recall increases by 0\% and -0.1\%, 0.2\% and -0.1\% and 0.1\% and -0.1\% for SVM and KNN; and F1-measure increases by 0.2\% and 0.1\%, 0.4\% and 0.1\% and 0.3\% and 0.1\% for SVM and KNN when L-LDA, DL-LDA and LCA-based topic features are used in conjunction combination with lexical and punctuation features, respectively. Similar to results in Table \ref{tab:result_boundary}, NB and KNN demonstrate the lowest and moderate performance on the task of identifying ``same segment'', respectively. \\
